### 2022-12-12 lmcs/introduction.tex
update Done

In the $\pi$-calculus literature, there have been several attempts at developing Curry-Howard correspondences between session-typed $\pi$-calculus and linear logic~\cite{girard87}: Caires and Pfenning's $\pi$DILL~\cite{cairespfenning10} corresponds to dual intuitionistic linear logic~\cite{barber96}, and Wadler's Classical Processes~\cite[CP]{wadler14} corresponds to classical linear logic~\cite[CLL]{girard87}. Both calculi guarantee deadlock freedom, which they achieve by the combination of binary session types with a restriction on the way processes can be connected by channels: two processes can share at most one channel, and, more generally, information transmitted between any two processes must pass through one unique series of channels and intermediate processes. We refer to such processes as \emph{tree-structured}, because the \emph{communication graph}---where the vertices are ready processes and two vertices are connected by an edge if and only if the corresponding processes share a channel---is a tree. For $\pi$DILL, CP, and GV, tree-structure follows from a syntactic restriction: the combination of name restriction and parallel composition into a single syntactic construct, corresponding to the logical cut.

@@

Our aim is to develop a more expressive version of GV while maintaining deadlock freedom. While process calculi have their advantages, \eg, their succinctness compared to concurrent $\lambda$-calculi, we chose to work with GV for several reasons. In general, concurrent $\lambda$-calculi support higher-order functions, and have a capability for abstraction not usually present in process calculi. Within a concurrent $\lambda$-calculus, one can derive extensions of the communication capabilities of the language via well-understood extensions of the functional fragment, \eg, we can derive internal/external choice from sum types. Concurrent $\lambda$-calculi maintain a clear separation between the program which the user writes and the configurations which represent the state of the system as it evaluates the program. However, our main motivation is that results obtained for $\lambda$-calculi transfer more easily to real-world functional programming languages. Case in point: we easily adapted the type system of PGV to Linear Haskell~\cite{bernardyboespflug18}, which gives us a library for deadlock-free session-typed programming~\cite{kokkedardha21hs}.
The benefit of working specifically with GV, as opposed to other concurrent $\lambda$-calculi, is its relation to CP~\cite{wadler14}, and its formal properties, including deadlock freedom.

>>>>

In the $\pi$-calculus literature, there have been several developments of Curry-Howard correspondences between session-typed $\pi$-calculus and linear logic~\cite{girard87}: Caires and Pfenning's $\pi$DILL~\cite{cairespfenning10} corresponds to dual intuitionistic linear logic~\cite{barber96}, and Wadler's Classical Processes~\cite[CP]{wadler14} corresponds to classical linear logic~\cite[CLL]{girard87}. Both calculi guarantee deadlock freedom, which they achieve by the combination of binary session types with a restriction on the way processes can be connected by channels: two processes can share at most one channel, and, more generally, information transmitted between any two processes must pass through one unique series of channels and intermediate processes. We refer to such processes as \emph{tree-structured}, because the \emph{communication graph}---where the vertices are ready processes and two vertices are connected by an edge if and only if the corresponding processes share a channel---is a tree. For $\pi$DILL, CP, and GV, tree-structure follows from a syntactic restriction: the combination of name restriction and parallel composition into a single syntactic construct, corresponding to the logical cut.

@@

Our aim is to develop a more expressive version of GV while maintaining deadlock freedom. While process calculi have their advantages, \eg, their succinctness compared to concurrent $\lambda$-calculi, we chose to work with GV for several reasons. In general, concurrent $\lambda$-calculi support higher-order functions, and have a capability for abstraction not usually present in process calculi. Within a concurrent $\lambda$-calculus, one can derive extensions of the communication capabilities of the language via well-understood extensions of the functional fragment, \eg, we can derive internal/external choice from sum types. Concurrent $\lambda$-calculi, among other languages, maintain a clear separation between the program which the user writes and the configurations which represent the state of the system as it evaluates the program. However, our main motivation is that results obtained for $\lambda$-calculi transfer more easily to real-world functional programming languages. Case in point: we easily adapted the type system of PGV to Linear Haskell~\cite{bernardyboespflug18}, which gives us a library for deadlock-free session-typed programming~\cite{kokkedardha21hs}.
The benefit of working specifically with GV, as opposed to other concurrent $\lambda$-calculi, is its relation to CP~\cite{wadler14}, and its formal properties, including deadlock freedom.
